---
title: "Tutorial 2"
author: Matthias Nijman - S4667999
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
---

## 1

### 1A-I

Throwing one on every throw from the 10th throw onward

### 1A-II

Throwing one on any throw after the 5th

### 1A-III

Throwing anything but one on the 4th throw and then throwing one every
time afterwards.
<span style="color:red">*Anything but one on the first 4. See the cap not cup.*</span>

### 1B

$$
\left( \bigcup_{i>n} \varepsilon_i \right)
\;\cap\;
\left( \bigcap_{i \le n} \varepsilon_i^{c} \right)
$$

## 2

10 balls, 6 white, 4 red. R = red Probability of red on pick 2 given red
on pick 1 = $(R_2|R_1)$ After the first pick we have 9 balls, of which 3
are red. Thus $P(R_2|R_1) = \frac{3}{9}$.
<span style="color:red">*It should be $P(R_b|R_1 \cup R_2) = \frac{P(R_b)}{P(R_1) + P(R2) - P(R_b)} =\frac{1}{5}$. Mind that $P(R_b) = P(R_1) \cap P(R_2) = P(R_1) \times P(R_2)$*</span>

## 3

$B_1$ has 8 red, 5 blue, total 13 $B_2$ has 6 red, 4 blue, total 10 We
roll a fair die, 1 = $B_1$, other = $B_2$

### 3A

$P(R) = P(R|B_1)P(B_1) + P(R|B_2)P(B_2)$\
$P(B_1) = \frac{1}{6}$\
$P(B_2) = \frac{5}{6}$\
$P(R|B_1) = \frac{8}{13}$\
$P(R|B_2) = \frac{6}{10}$\
$P(R) = \frac{8}{13}\frac{1}{6} + \frac{6}{10}\frac{5}{6}$

```{r}
library(MASS)
p = (8/13*(1/6)) + (6/10*5/6)
fractions(p)
```

### 3B

After drawing from a bag we keep drawing from that bag.
$P(R_1 \& R_2) = (\frac{8}{13})^2 \frac{1}{6} + (\frac{6}{10})^2 \frac{5}{6}$

```{r}
library(MASS)
p = ((8/13)^2*(1/6)) + ((6/10)^2*5/6)
fractions(p)
```

### 3C

We are drawing with replacement, thus the chances remain the same after
each draw. Drawing red twice is $P(R)^2$ thus $(\frac{47}{78})^2$

```{r}
fractions((47/78)^2)
(47/78)^2
```

## 4
### 4A
```{r}
sample(x=c("H","T"), size=10, replace=TRUE)
```
We got 5 H and 5 T, The ratio of H is $\frac{5}{10} = \frac{1}{2}$

### 4B
```{r}
calculate_head_ratio <- function(n) {
  tosses <- sample(x=c("H","T"), size=n, replace=TRUE)
  count <- 0
  for (x in tosses) {
    if (x == "H") {
      count <- count +1
    }
  }
  return (count/length(tosses))
}

```

### 4C
```{r}
results <- sapply(1:10000, calculate_head_ratio)
# Added: type='l' and abline(h=0.5,col='red',lwd=2)
plot(results, type='l', main="Ratio of Heads in a coin toss", xlab="Tosses", ylab="Heads ratio")
abline(h=0.5,col='red',lwd=2)

```

The plot shows that as the number of throws goes up, the ratio gets closer to 0.5

### 5A
We will use $B$ for buying and $C$ for customer, and then 1, 2 and 3 for the category.\
$P(B) = P(B|C_1)+P(B|C_2)+P(B|C_3)$\
We know that $P(B|C_1)=0.4$, $P(B|C_2)=0.13$, $P(B|C_3)=0.08$, thus $P(B) = 0.4+0.13+0.08=0.61$.
<span style="color:red">*Law of total prob: $P (ε) = P (ε | C1)P (C1) + P (ε | C2)P (C2) + P (ε | C3)P (C3)$*</span>


### 5B
$P(C_3) = 0.25$
Using Bayes' Formula, we know that:\
$P(C_3|B) = \frac{P(B|C_3)P(C_3)}{P(B)} = \frac{P(0.08)P(0.25)}{P(0.61)}$
```{r}
0.08*0.25/0.61
fractions(0.08*0.25/0.61)
```

$P(C_3|B) = 0.03278689 = \frac{2}{61}$

### 6A
The probability the student answers correctly on a question they knew the answer to. In this case 1.

### 6B
The probability a student knew the question given they answered it correctly. \

### 6C
$P(R) = \frac{1}{N}+0.7$\
$P(K|R) > \frac{P(R|K)P(K)}{P(R)}$ \
$0.99 > \frac{1\times0.7}{\frac{1}{N}+0.7}$\
$\frac{1}{N} < \frac{0.7}{0.99} - 0.7$\
$N < 141.4$\

<span style="color:red">*$\frac{7N}{7N+3}> 0.99$ so N > 43*</span>


## 7
### 7A
They are independent if $P(\epsilon_1\cap\epsilon_2)=P(\epsilon_1)P(\epsilon_2)$.
We know $P(\epsilon_1\cap\epsilon_2)=0.65$ and $P(\epsilon_1)P(\epsilon_2)=0.5\times0.2=0.1$.
This means they are not independent.

### 7B
$P(\epsilon_1\cap\epsilon_2)\geq P(\epsilon_1) + P(\epsilon_2) - 1$\
$P(\epsilon_1) + P(\epsilon_2) - 1 = 0.8 + 0.4 - 1 = 0.2$\
The statement is correct.

### 7C
Events are independent if $P(\epsilon_1\cap\epsilon_2)=P(\epsilon_1)P(\epsilon_2)$.\
And $P(\epsilon_1\cap\epsilon_2)= P(\epsilon_1) + P(\epsilon_2) - 1$.\

For $\emptyset$:\
$P(\emptyset\cap\epsilon)= P(\emptyset) + P(\epsilon) - 1$\
$0 + P(\epsilon) - 1 \leq 0$ this intersection cannot be smaller than 0 in reality, so it will equal 0.\
$P(\emptyset)P(\epsilon) = 0\times P(\epsilon) = 0$.\
$P(\emptyset\cap\epsilon)=P(\emptyset)P(\epsilon)$ is satisfied, thus $\emptyset$ is independent from any event.

For $\Omega$:\
$P(\Omega\cap\epsilon)= P(\Omega) + P(\epsilon) - 1$\
$1 + P(\epsilon) - 1 = P(\epsilon)$\
$P(\Omega)P(\epsilon) = 1\times P(\epsilon) = P(\epsilon)$.\
$P(\Omega\cap\epsilon)=P(\Omega)P(\epsilon)$ is satisfied, thus $\Omega$ is independent from any event.

### 7D
All three are independent, so $P(\epsilon_1\cap\epsilon_2\cap\epsilon_3) = P(\epsilon_1) P(\epsilon_2) P(\epsilon_3)$.\
That means we can say that $P((\epsilon_1\cap\epsilon_2)\cap\epsilon_3)=(P(\epsilon_1)P(\epsilon_2))P(\epsilon_3)$.\
Which means they are independent.